---
title: "meeple-explore.Rmd"
author: "mikec964"
date: "10/17/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)

colors.neon4 <- c("#a763ff", "#ffd063", "#bbff63", "#ff63bb")
source("scripts/load_and_wrangle.R")
```


## Explore the customer
* What's the distribution of his/her ratings?
* Distribution of game attributes: published, #players, playtime
* Distribution of game tags: categories, mechanics
* Correlation between ratings and attributes

```{r echo=FALSE, message=TRUE, warning=TRUE}
ggplot(games.details, aes(rating)) +
  ggtitle("Customer Ratings of Collection") +
  geom_bar(fill=colors.neon4[1])

ggplot(games.details, aes(x=yearpublished, y=rating)) +
  ggtitle("Cutomer's Rating vs. Publish Date") +
  geom_jitter(alpha=0.5, color=colors.neon4[1])

ggplot(games.details, aes(x=playingtime, y=rating)) +
  ggtitle("Cutomer's Rating vs. Playing Time") +
  geom_jitter(alpha=0.5, color=colors.neon4[1])

ggplot(games.details, aes(x=maxplayers, y=rating)) +
  ggtitle("Cutomer's Rating vs. Max Players") +
  geom_jitter(alpha=0.5, color=colors.neon4[1])

```


```{r echo=FALSE, message=TRUE, warning=TRUE}
# Which mechanics do I prefer? (# of games per, ratings of games per)
# Which categories do I prefer?

# Per mechanic, get the # of games and mean rating
mechanics.details <- data_frame(mechanic=games.mechanics,
                                count=0,
                                mean=0.0)
#mechanics.details$count <- as.integer(0)
#mechanics.details$mean <- as.numeric(0)
# for(m in g.mechanics$mechanics) {
#   g <- g.mechanics[g.mechanics$mechanics == m,]
#   r <- games.ratings[g$game.id, "rating"]
#   mechanics.details[mechanics.details$mechanic == m,]$count <- count(g)
#   mechanics.details[mechanics.details$mechanic == m,]$mean <- mean(r$rating)
# }
# mechanics.details$count <- as.integer(mechanics.details$count)
# mechanics.details$mean <- as.numeric(mechanics.details$mean)

categories.details <- data_frame(category=games.categories,
                                 count=0,
                                 mean=0)
# for(c in g.categories$categories) {
#   g <- g.categories[g.categories$categories == c,]
#   r <- games.ratings[g$game.id, "rating"]
#   categories.details[categories.details$category == c,]$count <- count(g)
#   categories.details[categories.details$category == c,]$mean <- mean(r$rating)
# }
# categories.details$count <- as.integer(categories.details$count)
# categories.details$mean <- as.numeric(categories.details$mean)
```

## Explore gamers
* How many games do gamers rate?
* What is distribution of their ratings?
* Distribution of game attributes: published, playtime, maxplayers

```{r echo=FALSE, message=TRUE, warning=TRUE}
# From selected gamers, find min, max, avg number of games rated
# gamers.selected is list of gamers
# collection.selected is all the games they've rated

ggplot(gamers.most, aes(ratings)) +
  geom_density(fill=colors.neon4[2]) +
  ggtitle("Ratings per Gamer, 2nd-3rd quartile")

# Ratings of all adjacent gamers
ggplot(games.ratings, aes(rating)) +
  geom_bar(fill=colors.neon4[3]) +
  ggtitle("All Ratings of Collection")
  
# How many games does each adjacent gamers have in common with the customer?
#   In games.ratings, how many games does each gamer own?
c <- games.ratings %>% 
  group_by(gamer) %>% 
  count()
ggplot(c, aes(n)) +
  geom_density(fill=colors.neon4[1]) + 
  ggtitle("Games in common")

# safe bet: matrix completion
# fresh pick: graph network

```

## Explore games
* How well-known are the games (how many ratings each)?
* How well-regarded are the games (mean, sd, scatter of ratings)?
* When were the games published?
* Which game categories dominate?
* Which game mechanics dominate?
* Which game designers dominate?
* Which game publishers dominate?

