---
title: "meep-matcher.Rmd"
author: "mikec964"
date: "9/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
file.sources = list.files(c("scripts"),
                          pattern="*.R$", full.names=TRUE,
                          ignore.case=TRUE)
sapply(file.sources, source)
```

# Overview
Here's the plan, so far:

1. **Load the customer's collection.**  
Load the collection of games (and ratings) of the user. All users are gamers,
but the person we're building the recommendation for is the customer.
We'll build a Collections table with a row per game.

2. **Per game, load the ratings.**  
For each game in the customer's collection, get the gamers that also rated
it. Row per gamer/rating.

3. **Load collections of gamers with games in common**  
This can be a big list, so first apply a basic filter or take a sample.

For selected gamers, load their collections. Games
in their collections that are unknown to the customer are possible
recommendations.

4. **Explore the data**
Look at the customer's preferences, game attributes, and preferences of
adjacent gamers in order to suggest recommendation algorithms and weights.

5. **Rank gamers to make recommendations**


# Load the customer's collection
This is probably in the range of 10 to 200 games. 
Note that for some customers, the number of columns might be different.
If a customer has a game with wishlist=1, then for that row the column
"wishlistpriority" is added.

```{r echo=FALSE, message=FALSE, warning=TRUE}
customer <- "mikec"
collection.tbl <- GetGamerCollection(customer)
head(collection.tbl)
```


# Per game, load the ratings
For each game in the collection, we'll collect the data and ratings. The data
are things like number of players and year published. Ratings is a list of
potentially thousands of ratings between 1 and 10.

Sizing the data: Each game could have 10,000 or more ratings. The API delivers
100 ratings at a time, so it could take 100 queries to load all for a game. 
If the customer has 100 games, that's 10,000 queries!

I've written a simple cache system so that subsequent reads come from local
storage.

```{r echo=FALSE, message=FALSE, warning=TRUE}
sample.tbl <- collection.tbl
# sample.tbl <- sample_n(collection.tbl, 2)  # Uncomment to load fewer games
head(sample.tbl)
for(g in 1:dim(sample.tbl)[1]) {
  game.id <- sample.tbl[[g, "game.id"]]
  num.comments <- GetGameData(game.id)[[1, "comments"]]
  num.pages <- as.integer(((num.comments - 1) / 100) + 1)
  print(sprintf("%s of %s, loading %s comments for %s, %s",
                g, dim(sample.tbl)[1], num.comments, game.id, sample.tbl[[g, "game"]]))
  if(g == 1) {
    attrs <- GetGameData(game.id)
  } else {
    attrs <- bind_rows(attrs, GetGameData(game.id))
  }
  for(p in 1:num.pages) {
    if(g==1 && p == 1) {
      ratings.tbl <- GetGameRatings(game.id, page=p)
    } else {
      ratings.tbl <- bind_rows(ratings.tbl, GetGameRatings(game.id, page=p))
    }
  }
}
head(attrs)
head(ratings.tbl)

write_delim(ratings.tbl, "tables/ratings.tbl", delim = "\t", na = "NA", append = FALSE, col_names = TRUE)
write_delim(collection.tbl, "tables/collection.tsv", delim = "\t", na = "NA", append = FALSE, col_names = TRUE)
write_delim(sample.tbl, "tables/sample.tsv", delim = "\t", na = "NA", append = FALSE, col_names = TRUE)
write_delim(games.tbl, "tables/games.tsv", delim = "\t", na = "NA", append = FALSE, col_names = TRUE)

```


# Load collections of gamers with games in common
My collection of 70 games has a total of 1,349K ratings, from 154K gamers.
All of these gamers have at least one game in common with me. For some of
them we'll download their collections.

We should prioritize gamers by number of games in common with me.

```{r echo=FALSE, message=TRUE, warning=TRUE}
# Get unique gamers
# For now, we ignore every good idea above and just select 500 at random
adjacent_gamers <- unique(ratings.tbl$gamer)
fresh_data <- TRUE
if(fresh_data) {
  # Select new gamers, probably download data for them
  selected_gamers <- sample(adjacent_gamers, 500) # 154K adjacent, use only 500
  write_lines(selected_gamers, "tables/selected-gamers.tsv")
} else {
  # Use previously selected gamers, probably read cached data for them
  selected_gamers <- read_lines("tables/selected-gamers.tsv")
}

for(gamer in selected_gamers) {
  print(sprintf("loading: %s", gamer))
  gamer_collection.tbl <- GetGamerCollection(gamer)
    if(gamer == selected_gamers[1]) {
      collections.tbl <- gamer_collection.tbl
    } else {
      collections.tbl <- bind_rows(collections.tbl, gamer_collection.tbl)
    }
}
```

# Explore the data
## Explore the customer
* What's the distribution of his/her ratings?
* Distribution of game attributes: published, categories, mechanics, #players, playtime
* Correlation between ratings and attributes

```{r echo=FALSE, message=TRUE, warning=TRUE}
rated.games <- collection.tbl[!is.na(collection.tbl$rating),]
ggplot(rated.games, aes(rating)) +
  geom_bar(fill="#6ec75c")

# Do I like new games better?
rated2 <- left_join(rated.games, published, by="game.id")
published.plot <- ggplot(rated2, aes(x=yearpublished, y=rating)) +
  ggtitle("Rating vs. Publish Date")
published.plot + geom_point(alpha=0.5, color="#c46c5b")
published.plot + geom_jitter(alpha=0.5, color="#c45b9f")

# Which mechanics do I prefer? (# of games per, ratings of games per)
mechanics <- attrs[attrs$key == "boardgamemechanic",]
mechanics.unique <- unique(mechanics$value)

# Which categories do I prefer?
categories <- attrs[attrs$key == "boardgamecategory",]
categories.unique <- unique(categories$value)
```

## Explore gamers
* How many games do gamers rate?
* What is distribution of their ratings?
* Distribution of game attributes: published, categories, mechanics, #players, playtime

```{r echo=FALSE, message=TRUE, warning=TRUE}

```

## Explore games
* How well-known are the games (how many ratings each)?
* How well-regarded are the games (mean, sd, scatter of ratings)?
* When were the games published?
* Which game categories dominate?
* Which game mechanics dominate?
* Which game designers dominate?
* Which game publishers dominate?


# Rank gamers to make recommendations
The general idea is:  
  * For the games we have rated, find gamers who have rated them similarly  
  * The more games they agree with us on, the better
  * If we agree on a game that everyone agrees on, weigh that less
  * If we strongly agree, weigh that more
  * If we agree on category, weigh that more
  * If we agree on mechanics, weigh that more
  * If they haven't rated many games, weigh them less

```{r echo=FALSE, message=TRUE, warning=TRUE}

```


