---
title: "meep-matcher.Rmd"
author: "mikec964"
date: "10/15/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
library(tidyr)
file.sources <- list.files(c("scripts"),
                          pattern="*.R$", full.names=TRUE,
                          ignore.case=TRUE)
sapply(file.sources, source)

kFreshData <- FALSE
```

# Overview
This loads data from BGG's XML API and creates the following variables,
and TSV files for faster reloading in the future.

| variable            | count     | Description
| --------            |----------:| -----------
| customer            | 1         | user
| collection.customer | 160       | games in collection
| collection.sample   | 160       | customer games we'll use
| gamers.adjacent     | 158,705   | have games in common
| gamers.selected     | 500       | subset of adjacent_gamers
| collection.selected | 80,342    | games owned by selected_gamers + customer
| games.attrs         | 5954      | (tall) avg of 37 per game, 160 games
| games.ratings       | 1,385,920 | (tall) for games in collection


Here's the plan, so far:

1. **Load the customer's collection.**  
Load the collection of games (and ratings) of the user. All users are gamers,
but the person we're building the recommendation for is the customer.

2. **Per game, load the ratings.**  
For each game in the customer's collection, get the gamers (adjacent gamers)
that also rated it.

3. **Load collections of gamers with games in common**  
For selected adjacent gamers, load their collections. Games
in their collections that are unknown to the customer are possible
recommendations.

# Load the customer's collection
This is probably in the range of 10 to 200 games. 
Note that for some customers, the number of columns might be different.
If a customer has a game with wishlist=1, then for that row the column
"wishlistpriority" is added.

```{r echo=FALSE, message=FALSE, warning=TRUE}
customer <- "mikec"
if(kFreshData) {
  collection.customer <- GetGamerCollection(customer)
  write_tsv(collection.customer, "tables/collection-customer.tsv", 
            na = "NA", append = FALSE, col_names = TRUE)
} else {
  collection.customer <- read_tsv("tables/collection-customer.tsv")
}
head(collection.customer)
```


# Per game, load the ratings
For each game in the collection, we'll collect the data and ratings. The data
are things like number of players and year published. Ratings is a list of
potentially thousands of ratings between 1 and 10.

Sizing the data: Each game could have 10,000 or more ratings. The API delivers
100 ratings at a time, so it could take 100 queries to load all for a game. 
If the customer has 100 games, that's 10,000 queries!

I've written a simple cache system so that subsequent reads come from local
storage.

```{r echo=FALSE, message=FALSE, warning=TRUE}
# Select a subset of the collection
if(kFreshData) {
  collection.sample <- collection.customer
  # collection.sample <- sample_n(collection.customer, 2)  # load subset
  write_tsv(collection.sample, "tables/collection-sample.tsv", 
            na = "NA", append = FALSE, col_names = TRUE)
} else {
  collection.sample <- read_tsv("tables/collection-sample.tsv")
}
```

```{r echo=FALSE, message=FALSE, warning=TRUE}
# Load the attributes and ratings for selected games
if(kFreshData) {
  GetGameData(sort(collection.sample$game.id)) # defines games.attrs, games.ratings
  write_tsv(games.attrs, "tables/games-attrs.tsv",
            na = "NA", append = FALSE, col_names = TRUE)
  
  write_tsv(games.ratings, "tables/games-ratings.tsv",
            na = "NA", append = FALSE, col_names = TRUE)
} else {
  games.attrs <- read_tsv("tables/games-attrs.tsv")
  games.ratings <- read_tsv("tables/games-ratings.tsv")
}
```


# Load collections of gamers with games in common
My collection of 70 games has a total of 1,349K ratings, from 154K gamers.
All of these gamers have at least one game in common with me. For some of
them we'll download their collections.

We should prioritize gamers by number of games in common with me.

```{r echo=FALSE, message=FALSE, warning=TRUE}
# Get unique gamers
# For now, we ignore every good idea above and just select 500 at random
gamers.adjacent <- unique(games.ratings$gamer)
if(kFreshData) {
  # Select new gamers, probably download data for them
  gamers.selected <- sort(sample(gamers.adjacent, 1000)) # 154K adjacent, use only 500
  write_lines(gamers.selected, "tables/gamers-selected.tsv")
  collection.selected <- GetGamerCollection(gamers.selected)
  write_tsv(collection.selected, "tables/collection-selected.tsv",
            na = "NA", append = FALSE, col_names = TRUE)
} else {
  kDownloadCollections <- FALSE
  if(kDownloadCollections) {
    # Select from cached gamers
    gamers.cached <- str_extract(list.files(c("xmlcache/collection")),
                                 "(?<=username=)(.*?)(?=[.]xml$)")
    gamers.selected <- sort(sample(gamers.cached, 1000))
    write_lines(gamers.selected, "tables/gamers-selected.tsv")
    collection.selected <- GetGamerCollection(gamers.selected)
    write_tsv(collection.selected, "tables/collection-selected.tsv",
              na = "NA", append = FALSE, col_names = TRUE)
  } else {
    # Fastest, just load previously cached collection data
    gamers.selected <- read_lines("tables/gamers-selected.tsv")
    collection.selected <- read_tsv("tables/collection-selected.tsv")
  }
}
```

