---
title: "meep-matcher.Rmd"
author: "mikec964"
date: "7/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
Here's the plan, so far:
* Load the collection of games (and ratings) of the user/gamer, "customer". Row per game.
* For each game in the customer's collection, get the gamers that rated it. Row per gamer/rating.
* Add to table, some columns empty.
* For each of those gamers, load their collections (and ratings), extend and update table


# Load data for two users
We're loading two just to make sure this works on n+1 cases.

Note that number of columns might be different. This might happen if one user has some games on wishlist=1, then the column "wishlistpriority" is added.

```{r echo=FALSE, message=FALSE, warning=FALSE}
source("scripts/get_gamer_collection.R")
customer.df <- get_gamer_collection("mikec")
eg.df <- get_gamer_collection("educatedgravy")
head(customer.df)
head(eg.df)
```
# For each game in the customer collection, get the ratings
We'll append these rows to our customer collection table. Possible problems ahead:
* Each game could have 10,000 or more ratings...
  * 100 games is a million rows
  * Reading ratings 100 at a time (per API) means 10,000 queries!
* On the new rows the gamer, game, game_id, rating columns will be filled, but
the own, owned, wishlist columns will be empty until we load data for each gamer
* Since ratings are sorted best to worst, and delivered in pages of 100, we could:
  * Stop getting ratings below 5
  * Select some pages to sample

For now, we'll load the first page of (100) ratings.
```{r echo=FALSE, message=FALSE, warning=FALSE}
source("scripts/get_game_data.R")
game_list <- customer.df[, "game_id"]
game_ratings <- get_game_data(game_list[1])
head(game_ratings)
```

# Combine customer game ratings with other gamer's ratings
