---
title: "Meeple Matcher Statistics"
author: "mikec964"
date: "7/20/2018"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(scales)
library(tidyr)

source("scripts/restore_data.R")
```

# Compare collection sizes
The project goal is to make game recommendations for the customer.

How does the collection size of the customer compare to other gamers?
In this case we loaded the collections for all gamers who had at least one game
in common with the customer.

This resulted in loading 158K gamers who had rated 1.385 million games. The
customer has 160 games in their collection, though only 60 are rated.
Collections were loaded for 1000 random gamers.
```{r echo=FALSE, message=TRUE, warning=TRUE}
collection_size_tbl <- gamer_collections_tall %>%
  group_by(gamer) %>%
  count(sort=TRUE)
ggplot(collection_size_tbl, aes(n)) +
  geom_histogram(binwidth=10) +
  labs(title="Sizes of Game Collections (of 1000 gamers)", x="Games", y="No. of Gamers")
```

```{r echo=FALSE, message=TRUE, warning=TRUE}
# q10_qty_tbl <- gamer_collections_tall %>%
#   # 10 quantiles by #games per gamer
#   count(gamer) %>%
#   arrange(n) %>%
#   mutate(quant = rep(1:10, each=ceiling(length(gamer)/10))) %>%
#   # mean collection size per quintile
#   group_by(quant) %>%
#   summarize(games_mu = as.integer(mean(n, na.rm=TRUE)),
#             games_sd = as.integer(sd(n, na.rm=TRUE)),
#             games_max = max(n, na.rm=TRUE))
# ggplot(data=q10_qty_tbl, mapping=aes(x=quant, y=games_mu)) +
#   scale_x_continuous(breaks=seq(1,10)) +
#   scale_y_continuous(breaks=seq(0,1300, by=100)) +
#   labs(title="Games Rated per Gamer", x="Quantile", y="Mean") +
#   geom_col() +
#   geom_text(aes(x=quant, y=games_mu + 60, label=games_mu))
```

# Compare customer ratings to most ratings
These charts show that ratings from the customer and other gamers both follow
a normal distribution, slightly skewed left.
```{r echo=FALSE, message=TRUE, warning=TRUE}
rating_props_tbl <- game_ratings_tall %>%
  mutate(is_customer = (gamer == customer)) %>%
  group_by(rating, is_customer) %>%
  summarize(n = sum(rating)) %>%
  group_by(is_customer) %>%
  mutate(prop = n/sum(n)) %>% # proportion of customer or non-customer votes
  filter(rating > 0)

ggplot(data=rating_props_tbl, mapping=aes(x=rating, y=prop, fill=is_customer)) +
  labs(title="Game Ratings (1-10)", x=NULL, y="Percentage") +
  scale_fill_discrete(name="Gamer", labels=c("By others", "By customer")) +
  scale_x_continuous(breaks=seq(1,10)) +
  scale_y_continuous(labels = percent_format()) +
  geom_col(position="dodge")
# note: ratings are screwed left
```

# Rating per game
Next, we learn a little more about what kinds of games the customer likes.
Games have *mechanics*, like card-drafting, dice-rolling, and tile-laying. They
also have *categories* like science-fiction, co-operative, and family.

First we look at which mechanics are most common in the customer's collection.
```{r echo=FALSE, message=TRUE, warning=TRUE}
# calc mean rating per game
game_ratings_tbl <- gamer_collections_tall %>%
  group_by(game.id) %>%
  summarize(rating_mean = mean(rating, na.rm=TRUE),
            rating_sd = sd(rating, na.rm=TRUE)) %>%
  filter(!is.nan(rating_mean)) %>%
  filter(!is.na(rating_sd))
# result: 27k games (16K rated, 8K with >1 rating)

# Games per mechanic ----------------------------------------------------
# game_attrs is tall, make it tidy with observation per game
game_mechs_tall <- game_attrs_tall %>%
  # keep only the mechanics data
  group_by(game.id) %>%
  filter(key == "boardgamemechanic") %>%
  select(-one_of("key")) %>%
  rename(mech=value)

# make each mechanic an observation with vars: n games
mech_counts_tbl <- game_mechs_tall %>%
  group_by(mech) %>%
  summarize(n=n()) %>%
  arrange(n) %>%
  # top 20 mechanics
  top_n(20, n)

ggplot(mech_counts_tbl, aes(reorder(mech, n), n)) +
  labs(title="Games per Mechanic", x="top 20 mechanics", y=NULL) +
  geom_col() +
  coord_flip()
```

It should be more interesting to know which mechanics are in the games that the
customer likes. This chart shows that the customer often buys games with
variable player powers and rates them higher. They less often buy games with
player elimination and they rate them lower.

```{r echo=FALSE, message=TRUE, warning=TRUE}
# Ratings per mech ---------------------------------------------------
game_mechs_tall$rating <- sapply(game_mechs_tall$game.id, function(x){
  as.numeric(game_ratings_tbl[game_ratings_tbl$game.id == x, "rating_mean"])
})

mech_ratings_tbl <- game_mechs_tall %>%
  # cols will be games, values will be game ratings
  group_by(mech) %>%
  summarize(rating_mu = mean(rating, na.rm=TRUE))

ggplot(mech_ratings_tbl, aes(mech, rating_mu)) +
  labs(title="Ratings per Mechanic", x=NULL, y="Mean Rating") +
  scale_y_discrete(breaks=seq(1, 10, by=0.5)) +
  geom_col() +
  coord_flip()

mech20_tbl <- mech_counts_tbl %>%
  left_join(mech_ratings_tbl, by="mech")
ggplot(mech20_tbl, aes(mech, n, fill=rating_mu)) +
  labs(title="Game Mechanics and Ratings",
       x="top 20 mechanics", y="no. of games") +
  geom_col() +
  coord_flip()
```

# Games per category
I look at game categories the same way.

```{r echo=FALSE, message=TRUE, warning=TRUE}
# game_attrs is tall, make it tidy with observation per game
game_gcats_tall <- game_attrs_tall %>%
  # keep only the mechanics data
  group_by(game.id) %>%
  filter(key == "boardgamecategory") %>%
  select(-one_of("key")) %>%
  rename(gcat=value)

# make each gcategory an observation with vars: n games
gcat_counts_tbl <- game_gcats_tall %>%
  group_by(gcat) %>%
  summarize(n=n()) %>%
  arrange(n) %>%
  top_n(20, n)

# this orders the plot to match the table
ggplot(gcat_counts_tbl, aes(reorder(gcat, n), n)) +
  labs(title="Games per Category", x="top 20 categories", y=NULL) +
  geom_col() +
  coord_flip()


# Ratings per category ----------------------------------------
game_gcats_tall$rating <- sapply(game_gcats_tall$game.id, function(x){
  as.numeric(game_ratings_tbl[game_ratings_tbl$game.id == x, "rating_mean"])
})

gcat_ratings_tbl <- game_gcats_tall %>%
  # cols will be games, values will be game ratings
  group_by(gcat) %>%
  summarize(rating_mu = mean(rating, na.rm=TRUE))

ggplot(gcat_ratings_tbl, aes(gcat, rating_mu)) +
  labs(title="Ratings per Category", x=NULL, y="Mean Rating") +
  scale_y_discrete(breaks=seq(1, 10, by=0.5)) +
  geom_col() +
  coord_flip()

gcat20_tbl <- gcat_counts_tbl %>%
  left_join(gcat_ratings_tbl, by="gcat")
ggplot(gcat20_tbl, aes(gcat, n, fill=rating_mu)) +
  labs(title="Game Categories and Ratings",
       x="top 20 categories", y="no. of games") +
  geom_col() +
  coord_flip()
```
The customer has mostly science fiction games and rates them higher, and few
movie-themed games rated lower. There are counter-examples, too: Miniatures
games are purchased infrequently but highly rated.

This might be an opportunity to recommend a type of game to play.
