---
title: "Meeple Matcher Statistics"
author: "mikec964"
date: "7/20/2018"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list=ls())
library(dplyr)
library(ggplot2)
library(ggrepel)
library(scales)
library(tidyr)

source("scripts/restore_data.R")
```

# Compare collection sizes
The project goal is to make game recommendations for the customer.

How does the collection size of the customer compare to other gamers? Over
158,000 gamers had at least one game in common with the customer. I selected
1000 of those gamers at random and loaded their collections.

```{r collection size histogram, echo=FALSE, message=TRUE, warning=TRUE}
n_collection_cust <- nrow(subset(gamer_collections_tall,
                                 gamer == customer))
n_rated_cust <- nrow(subset(gamer_collections_tall,
                            gamer == customer & !is.na(rating)))
n_collection_all <- nrow(subset(gamer_collections_tall,
                                gamer != customer))
n_rated_all <- nrow(subset(gamer_collections_tall,
                           gamer != customer & !is.na(rating)))
n_ratings <- nrow(bg_ratings_tall)

paste0("The customer has ",
       n_collection_cust, " games in their collection, ",
       "and has rated ", n_rated_cust, ".")
paste0("Of 1000 random gamers, they have ", 
       n_collection_all, " games collected, ",
       "and have rated ", n_rated_all, ", for ", 
       n_ratings, " in total.")

collection_size_tbl <- gamer_collections_tall %>%
  group_by(gamer) %>%
  count(sort = TRUE) %>%
  arrange(n)
ggplot(collection_size_tbl, aes(n)) +
  geom_histogram(binwidth = 10) +
  labs(
    title = "Sizes of Game Collections (of 1000 gamers)",
    x = "Games", y = "No. of Gamers"
  )
ggplot(collection_size_tbl[1:800, ], aes(n)) +
  geom_histogram(binwidth = 10) +
  labs(
    title = "Sizes of Game Collections (of bottom 800 gamers)",
    x = "Games", y = "No. of Gamers"
  ) +
  geom_segment(aes(x=n_rated_cust, xend=n_rated_cust, 
                   y=0, yend=60), color="turquoise") +
  geom_label(aes(x=250, y=50), label="Customer's collection in green")

```

```{r collection size quintiles, echo=FALSE, message=TRUE, warning=TRUE}
# q10_qty_tbl <- gamer_collections_tall %>%
#   # 10 quantiles by #games per gamer
#   count(gamer) %>%
#   arrange(n) %>%
#   mutate(quant = rep(1:10, each=ceiling(length(gamer)/10))) %>%
#   # mean collection size per quintile
#   group_by(quant) %>%
#   summarize(games_mu = as.integer(mean(n, na.rm=TRUE)),
#             games_sd = as.integer(sd(n, na.rm=TRUE)),
#             games_max = max(n, na.rm=TRUE))
# ggplot(data=q10_qty_tbl, mapping=aes(x=quant, y=games_mu)) +
#   scale_x_continuous(breaks=seq(1,10)) +
#   scale_y_continuous(breaks=seq(0,1300, by=100)) +
#   labs(title="Games Rated per Gamer", x="Quantile", y="Mean") +
#   geom_col() +
#   geom_text(aes(x=quant, y=games_mu + 60, label=games_mu))
```

# Compare customer ratings to most ratings
These charts show that ratings from the customer and other gamers both follow
a normal distribution, slightly skewed left.

```{r Bar chart of games per rating, echo=FALSE, message=TRUE, warning=TRUE}
rating_props_tbl <- bg_ratings_tall %>%
  mutate(is_customer = (gamer == customer)) %>%
  group_by(rating, is_customer) %>%
  summarize(n = sum(rating)) %>%
  group_by(is_customer) %>%
  mutate(prop = n / sum(n)) %>% # proportion of customer or non-customer votes
  filter(rating > 0)

ggplot(rating_props_tbl, aes(x = rating, y = prop, fill = is_customer)) +
  labs(title = "Game Ratings (1-10)", x = NULL, y = "Percentage") +
  scale_fill_discrete(name = "Gamer", labels = c("By others", "By customer")) +
  scale_x_continuous(breaks = seq(1, 10)) +
  scale_y_continuous(labels = percent_format()) +
  geom_col(position = "dodge")
# note: ratings are screwed left
```

```{r Calc ratings per game, echo=FALSE, message=TRUE, warning=TRUE}
# Also calc rating per game by customer and by others
bg_ratings_tbl <- gamer_collections_tall %>%
  filter(!is.na(rating)) %>%
  select(game.id, gamer, rating) %>%
  mutate(is_cust = (gamer == customer),
         c_rating = is_cust * rating,
         c_rating = na_if(c_rating, 0),
         a_rating  = (!is_cust) * rating,
         a_rating  = na_if(a_rating, 0))
bg_ratings_tbl <- bg_ratings_tbl %>%
  group_by(game.id) %>%
  summarize(n_ratings = n(),
            rating_by_cust = sum(c_rating, na.rm = TRUE),
            rating_by_cust = na_if(rating_by_cust, 0),
            ratings_by_all  = mean(a_rating, na.rm = TRUE))
# result: Of 27k games, 16K rated, 8K with >1 rating

```


# Rating per game attributes
Next, we learn a little more about what kinds of games the customer likes.
Games have *mechanics*, like card-drafting, dice-rolling, and tile-laying. They
also have *categories* like science-fiction, co-operative, and family.

First we look at which mechanics are most common in the customer's collection.

```{r Plot games per mechanic, echo=FALSE, message=TRUE, warning=TRUE}
#' Get a list of values for an attribute
#'
#' @param bg_attr "boardgamemechanic" or "boardgamecategory"
#' @param bg_attrs_tall A tall table of game.id and key/value pairs.
#' @param bg_ratings_tbl A table of games with customer and overall ratings.
#' @return A table of values and count of games with that value.
get_attr_stats <- function(bg_attr, 
                           attrs_tall = bg_attrs_tall, 
                           ratings_tbl = bg_ratings_tbl) {
  attr_stats <- attrs_tall %>%
    # tall with game.id * attribute (mechanic) rows,
    #     n (count of ratings), cust_rating, all_ratings
    group_by(game.id) %>%
    filter(key == bg_attr) %>%
    select(-one_of("key")) %>%
    inner_join(ratings_tbl, by = "game.id")
  ttl_cust_games <- sum(!is.na(bg_ratings_tbl$rating_by_cust))
  ttl_all_games <- sum(!is.na(bg_ratings_tbl$ratings_by_all))
  
  attr_stats2 <- attr_stats %>%
    # table of attribute values (mechanics),
    #   n_games (per attr value), and weighted ratings
    group_by(value) %>%
    summarize(n_games = n(),
              n_cust_games = sum(!is.na(rating_by_cust)),
              ratings_total = sum(n_ratings),
              cust_rating_wt = mean(rating_by_cust, na.rm = TRUE),
              all_ratings_wt = sum(ratings_by_all * n_ratings) 
                                / sum(n_ratings)) %>%
    arrange(desc(cust_rating_wt))
  
  attr_stats3 <- attr_stats2 %>%
    mutate(cust_ownership = n_cust_games / ttl_cust_games) %>%
    mutate(all_ownership = n_games / ttl_all_games) %>%
    mutate(own_ratio = cust_ownership / all_ownership)
}

mech_ratings_tbl <- get_attr_stats("boardgamemechanic")
ggplot(mech_ratings_tbl[1:25, ], aes(reorder(value, n_games), n_games)) +
  labs(title="Games per Mechanic", x="Top 20 Mechanics", y=NULL) +
  geom_col(fill="blue") +
  coord_flip()
```

For each mechanic, this chart shows the difference between the
customer and the overall rating.

```{r Plot ratings per mech, echo=FALSE, message=TRUE, warning=TRUE}
mech_ratings_tbl$pref <- mech_ratings_tbl$cust_rating_wt - mech_ratings_tbl$all_ratings_wt
ggplot(mech_ratings_tbl[1:25, ], aes(reorder(value, pref), pref)) +
  labs(title="Customer's Preferred Mechanics", x="Mechanic", 
       y="Rating by Customer - Rating by Others") +
  theme(legend.position="bottom") +
  geom_point() +
  geom_segment(aes(x=1, xend=25, y=0, yend=0), 
               color="blue") +
  coord_flip()
```

This helps identify mechanics that the customer likes and seeks out. Some
mechanics are more common in games than others. For each mechanic, the ratio
of games the customer owns with that mechanic against the customer's rating
of games with that mechanic.

We can see that the customer has many games with variable player powers and
hand management, and has few games with storytelling, and rates it lower.

```{r Plot ownership per mech, echo=FALSE, message=TRUE, warning=FALSE}
ggplot(mech_ratings_tbl, aes(cust_rating_wt, n_cust_games, label=value)) +
  labs(title="Customer Preference and Ownership per Mechanic", 
       x="Customer Rating", 
       y="Ratio of Customer to Overall Ownership") +
  geom_point() +
  geom_text_repel(size=2)
```

# Games per category
I look at game categories the same way. Which categories are often selected?

```{r games per category, echo=FALSE, message=TRUE, warning=TRUE}
gcat_ratings_tbl <- get_attr_stats("boardgamecategory")
ggplot(gcat_ratings_tbl[1:25,], aes(reorder(value, n_games), n_games)) +
  labs(title="Games per Category", x="Top 20 Categories", y=NULL) +
  geom_col(fill="purple") +
  coord_flip()
```

```{r Plot ratings per category, echo=FALSE, message=TRUE, warning=TRUE}
gcat_ratings_tbl$pref <- gcat_ratings_tbl$cust_rating_wt - gcat_ratings_tbl$all_ratings_wt
ggplot(gcat_ratings_tbl[1:25, ], aes(reorder(value, pref), pref)) +
  labs(title="Customer's Preferred Categories", x="Category", 
       y="Rating by Customer - Rating by Others") +
  theme(legend.position="bottom") +
  geom_point() +
  geom_segment(aes(x=1, xend=25, y=0, yend=0),
               color="purple") +
  coord_flip()
```

We see that the customer tends to buy card games and science fiction games, but
rates space exploration much higher.

```{r Plot ownership per category, echo=FALSE, message=TRUE, warning=FALSE}
ggplot(gcat_ratings_tbl, aes(cust_rating_wt, cust_ownership, label=value)) +
  labs(title="Customer Ownership vs Overall Ownership", 
       x="Customer Rating", 
       y="Ratio of Category Within Games") +
  geom_point() +
  geom_text_repel(size=2)
```

Next steps might be to identify groups of categories and mechanics based on
rating and ownership and games.

This might be an opportunity to recommend games based on attributes as well as
by ratings; a hybrid system.

