---
title: "Meeple Matcher Statistics"
author: "mikec964"
date: "7/20/2018"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list=ls())
library(dplyr)
library(ggplot2)
library(ggrepel)
library(scales)
library(tidyr)

source("scripts/restore_data.R")
```

# Compare collection sizes
The project goal is to make game recommendations for the customer.

How does the collection size of the customer compare to other gamers? Over
158,000 gamers had at least one game in common with the customer. I selected
1000 of those gamers at random and loaded their collections.

```{r collection size histogram, echo=FALSE, message=TRUE, warning=TRUE}
n_collection_cust <- nrow(subset(gamer_collections_tall,
                                 gamer == customer))
n_rated_cust <- nrow(subset(gamer_collections_tall,
                            gamer == customer & !is.na(rating)))
n_collection_all <- nrow(subset(gamer_collections_tall,
                                gamer != customer))
n_rated_all <- nrow(subset(gamer_collections_tall,
                           gamer != customer & !is.na(rating)))
n_ratings <- nrow(bg_ratings_tall)

paste0("The customer has ",
       n_collection_cust, " games in their collection, ",
       "and has rated ", n_rated_cust, ".")
paste0("Of 1000 random gamers, they have ", 
       n_collection_all, " games collected, ",
       "and have rated ", n_rated_all, ", for ", 
       n_ratings, " in total.")

collection_size_tbl <- gamer_collections_tall %>%
  group_by(gamer) %>%
  count(sort = TRUE) %>%
  arrange(n)
ggplot(collection_size_tbl, aes(n)) +
  geom_histogram(binwidth = 10) +
  labs(
    title = "Sizes of Game Collections (of 1000 gamers)",
    x = "Games", y = "No. of Gamers"
  )
ggplot(collection_size_tbl[1:800, ], aes(n)) +
  geom_histogram(binwidth = 10) +
  labs(
    title = "Sizes of Game Collections (of bottom 800 gamers)",
    x = "Games", y = "No. of Gamers"
  ) +
  geom_segment(aes(x=n_rated_cust, xend=n_rated_cust, 
                   y=0, yend=60), color="turquoise") +
  geom_label(aes(x=250, y=50), label="Customer's collection in green")
```

```{r collection size quintiles, echo=FALSE, message=TRUE, warning=TRUE}
# q10_qty_tbl <- gamer_collections_tall %>%
#   # 10 quantiles by #games per gamer
#   count(gamer) %>%
#   arrange(n) %>%
#   mutate(quant = rep(1:10, each=ceiling(length(gamer)/10))) %>%
#   # mean collection size per quintile
#   group_by(quant) %>%
#   summarize(games_mu = as.integer(mean(n, na.rm=TRUE)),
#             games_sd = as.integer(sd(n, na.rm=TRUE)),
#             games_max = max(n, na.rm=TRUE))
# ggplot(data=q10_qty_tbl, mapping=aes(x=quant, y=games_mu)) +
#   scale_x_continuous(breaks=seq(1,10)) +
#   scale_y_continuous(breaks=seq(0,1300, by=100)) +
#   labs(title="Games Rated per Gamer", x="Quantile", y="Mean") +
#   geom_col() +
#   geom_text(aes(x=quant, y=games_mu + 60, label=games_mu))
```

# Compare customer ratings to most ratings
These charts show that ratings from the customer and other gamers both follow
a normal distribution, slightly skewed left.

```{r Bar chart of games per rating, echo=FALSE, message=TRUE, warning=TRUE}
rating_props_tbl <- bg_ratings_tall %>%
  mutate(is_customer = (gamer == customer)) %>%
  group_by(rating, is_customer) %>%
  summarize(n = sum(rating)) %>%
  group_by(is_customer) %>%
  mutate(prop = n / sum(n)) %>% # proportion of customer or non-customer votes
  filter(rating > 0)

ggplot(rating_props_tbl, aes(x = rating, y = prop, fill = is_customer)) +
  labs(title = "Game Ratings (1-10)", x = NULL, y = "Percentage") +
  scale_fill_discrete(name = "Gamer", labels = c("By others", "By customer")) +
  scale_x_continuous(breaks = seq(1, 10)) +
  scale_y_continuous(labels = percent_format()) +
  geom_col(position = "dodge")
# note: ratings are screwed left
```

```{r Calc ratings, echo=FALSE, message=TRUE, warning=TRUE}
bg_ratings_tbl <- gamer_collections_tall %>%
  filter(!is.na(rating)) %>%
  select(game.id, gamer, rating) %>%
  mutate(is_cust = (gamer == customer),
         c_rating = is_cust * rating,
         c_rating = na_if(c_rating, 0),
         a_rating  = (!is_cust) * rating,
         a_rating  = na_if(a_rating, 0))
bg_ratings_tbl <- bg_ratings_tbl %>%
  group_by(game.id) %>%
  summarize(n_ratings = n(),
            rating_by_cust = sum(c_rating, na.rm = TRUE),
            rating_by_cust = na_if(rating_by_cust, 0),
            ratings_by_all  = mean(a_rating, na.rm = TRUE))
# result: Of 27k games, 16K rated, 8K with >1 rating

#' Get a list of values for an attribute
#'
#' @param bg_attr "boardgamemechanic" or "boardgamecategory"
#' @param bg_attrs_tall A tall table of game.id and key/value pairs.
#' @param bg_ratings_tbl A table of games with customer and overall ratings.
#' @return A table of values and count of games with that value.
get_attr_stats <- function(bg_attr, 
                           attrs_tall = bg_attrs_tall, 
                           ratings_tbl = bg_ratings_tbl) {
  attr_stats <- attrs_tall %>%
    # tall with game.id * attribute (mechanic) rows,
    #     n (count of ratings), cust_rating, all_ratings
    group_by(game.id) %>%
    filter(key == bg_attr) %>%
    select(-one_of("key")) %>%
    inner_join(ratings_tbl, by = "game.id")
  ttl_cust_games <- sum(!is.na(bg_ratings_tbl$rating_by_cust))
  ttl_all_games <- sum(!is.na(bg_ratings_tbl$ratings_by_all))
  
  attr_stats2 <- attr_stats %>%
    # table of attribute values (mechanics),
    #   n_games (per attr value), and weighted ratings
    group_by(value) %>%
    summarize(n_games = n(),
              n_cust_games = sum(!is.na(rating_by_cust)),
              ratings_total = sum(n_ratings),
              cust_rating_wt = mean(rating_by_cust, na.rm = TRUE),
              all_ratings_wt = sum(ratings_by_all * n_ratings) 
                                / sum(n_ratings)) %>%
    arrange(desc(cust_rating_wt))
  
  attr_stats3 <- attr_stats2 %>%
    mutate(cust_ownership = n_cust_games / ttl_cust_games) %>%
    mutate(all_ownership = n_games / ttl_all_games) %>%
    mutate(own_ratio = cust_ownership / all_ownership)
}
```


# Rating per game attributes
Next, we learn a little more about what kinds of games the customer likes.
Games have *mechanics*, like card-drafting, dice-rolling, and tile-laying. They
also have *categories* like science-fiction, co-operative, and family.

First we look at which mechanics are most common in the customer's collection.

```{r Plot games per mechanic, echo=FALSE, message=TRUE, warning=TRUE}
mech_ratings_tbl <- arrange(get_attr_stats("boardgamemechanic"),
                            desc(n_cust_games))
ggplot(mech_ratings_tbl[1:20, ], aes(reorder(value, n_games), n_games)) +
  labs(title="Customer's Games per Mechanic", x="Top 20 Mechanics", y=NULL) +
  geom_col(fill="blue") +
  coord_flip()
```

Since some mechanics are more common in games than others, we also look at the
customer's rating of games with each mechanic. I removed games rated under 6
from the chart. Some observations:
* Variable player powers is a top purchase and well-rated (6.7)
* Variable phase order rarely in purchases (not in top 20) but highly-rated (8.0)
* Cooperative play is both a top-5 purchase and highly-rated (6.9)

Mechanics at the bottom right of the table are in games the customer enjoyed
but does not have very many of, and might be a good place to look for fresh
recommendations.

Mechanics at the top right are in games the customer enjoyed and has often
purchased, so they might be safe bets for games that will be liked.

```{r Plot ownership per mech, echo=FALSE, message=TRUE, warning=FALSE}
mech_ratings_tbl <- subset(mech_ratings_tbl, cust_rating_wt >= 6)
ggplot(mech_ratings_tbl, aes(cust_rating_wt, own_ratio, label=value)) +
  labs(title="Game Mechanics in Collection", 
       x="Rating by Customer", 
       y="Games in Collection vs Overall") +
  geom_point(color="blue") +
  geom_text_repel(size=2)
```

For each mechanic, this chart shows the difference between the
customer's rating and the rating overall.
* Line drawing is far better rated by the customer than by others
* Variable player powers, often purchased and well-rated (6.73) by the customer,
is somewhat higher-rated by other players.

```{r Plot ratings per mech, echo=FALSE, message=TRUE, warning=TRUE}
mech_ratings_tbl$pref <- mech_ratings_tbl$cust_rating_wt - mech_ratings_tbl$all_ratings_wt
ggplot(mech_ratings_tbl[1:25, ], aes(reorder(value, pref), pref)) +
  labs(title="Customer's Preferred Mechanics", x="Mechanic", 
       y="Rating by Customer - Rating by Others") +
  theme(legend.position="bottom") +
  geom_point() +
  geom_segment(aes(x=1, xend=25, y=0, yend=0), 
               color="blue") +
  coord_flip()
```



# Games per category
Which game categories are common in the customer's collection?
* Science fiction and Space exploration are in the top 6; grouping might help.
* Dice rolling is a mechanic and Dice is a category. Categories might benefit
from manual grooming.


```{r games per category, echo=FALSE, message=TRUE, warning=TRUE}
gcat_ratings_tbl <- arrange(get_attr_stats("boardgamecategory"),
                            desc(n_cust_games))
ggplot(gcat_ratings_tbl[1:20,], aes(reorder(value, n_games), n_games)) +
  labs(title="Games per Category", x="Top 20 Categories", y=NULL) +
  geom_col(fill="purple") +
  coord_flip()
```

Here we show the number of games the customer owns with each category, and the
customer's rating of games with each category I removed games rated under
5.5. Some observations:
* We see that the customer tends to buy card games and science fiction games
* Space exploration (7.5) and Exploration games (7.3) are highly rated and
more commonly purchased.
* Nautical games are even higher rated (8.25) but less frequently purchased.

```{r Plot ownership per category, echo=FALSE, message=TRUE, warning=FALSE}
grt2 <- subset(gcat_ratings_tbl, cust_rating_wt >= 6)
ggplot(grt2, aes(cust_rating_wt, own_ratio, label=value)) +
  labs(title="Customer Preference and Ownership per Category", 
       x="Customer Rating", 
       y="Games in Collection vs Overall") +
  geom_point(color="purple") +
  geom_text_repel(size=2)
```




```{r Plot ratings per category, echo=FALSE, message=TRUE, warning=TRUE}
gcat_ratings_tbl$pref <- gcat_ratings_tbl$cust_rating_wt - gcat_ratings_tbl$all_ratings_wt
ggplot(gcat_ratings_tbl[1:20, ], aes(reorder(value, pref), pref)) +
  labs(title="Customer's Preferred Categories", x="Category", 
       y="Rating by Customer - Rating by Others") +
  theme(legend.position="bottom") +
  geom_point() +
  geom_segment(aes(x=1, xend=20, y=0, yend=0),
               color="purple") +
  coord_flip()
```



Next steps might be to identify groups of categories and mechanics based on
rating and ownership and games.

This might be an opportunity to recommend games based on attributes as well as
by ratings; a hybrid system.

